Word learning at the semantics pragmatics interface

### 2020-04-17 paper/supplementary_material.Rmd
finish edits through supplement; add a fully parameterized version of rsa math

```{r, include = F}
library(tidyverse)
library(rwebppl)
library(brms)
library(coda)
library(ggthemes)
library(readxl)
library(ggpubr)
library(stringr)
library(matrixStats)
library(data.table)
library(langcog)
library(ggimage)
library(BayesFactor)

@@

$$
P_{L}(r \mid u)\propto P_{S}(u \mid r) \cdot P(r \mid \rho_i) \\ 
P_{S}(u \mid r)\propto Informativity(u;r)^{\alpha_i} \\
Informativity(u; r) = P(r|u) \propto \mathcal{L}(u \mid \theta_{ij})
$$

@@

@@

@@

The first lesioned model takes in speaker informativeness and common ground sensitivity as well as general semantic knowledge (vocabulary development), but omits semantic knowledge that is specific to the familiar objects. We described above that the parameters for semantic knowledge are fitted via a hierarchical regression (mixed effects) model. In this model, there is an overall developmental trajectory for semantic knowledge (main effect) and then there is object-specific variation around this trajectory (random effects). The **no word knowledge model** takes in the overall trajectory for semantic knowledge---represented by $\mu^\theta_0$ and $\mu^\theta_1$---but ignores object-specific variation. Formally, the semantic knowledge for each item is the same: $\theta_{ij} = \mu^\theta_0 +  i \cdot \mu^\theta_1$. That is, the model assumes a listener whose mutual exclusivity inference does not vary depending on the particular familiar object but only depends on the average semantic knowledge.

@@

This model takes in object specific semantic knowledge and speaker informativeness but ignores common ground $\rho_i$. Thus, the prior distribution over objects in the model described above is uniform (e.g., [0.5, 0.5]). This corresponds to a listener who only focuses on the mutual exclusivity inference and ignores the common ground manipulation. As a consequence, the listener does not differentiate between the two common ground alignment conditions.

@@

The last lesioned model only takes in the common ground sensitivity. This corresponds to a listener who only focuses on common ground and ignores the identity of the objects on the tables as well as any inferences their semantic knowledge of the familiar objects license. The model predictions therefore correspond to the prior distribution over objects $P(r \mid \rho_i)$.

@@

We compared the models mentioned above in two ways. On the one hand, we used correlations between model predictions and the data. For this analysis, we binned the model predictions and the data by age in years and familiar object. Figure \@ref(fig:fig6) visualizes the correlation between model predictions and the data for all models. The results shows a very high correlation between the predictions of the **integration model** and the data in all age groups indicating that the model accurately captures the variation in the data. Correlations for the **integration model** were also higher compared to the other models considered. The correlation increased from 2- to 3-year-olds but then again dropped for 4-year-olds. The mis-match for the oldest age group is probably a consequence of the model making very extreme predictions in the congruent condition. This results from the fact that 4-year-olds show very high performance in the mututal exclusivity as well as the common ground task. When combined in the model, the two inferences amplify one another because we do not assume that there is any cost of integration. We made this choice because the parameter representing integration cost would have to be estimated based on the integration data of Experiment 3 itslef.

We additionally compared models based on the marginal likelihood of the data under each model -- the likelihood of the data averaging over ("marginalizing over") the prior distribution on parameters; the pair-wise ratio of marginal likelihoods for two models is known as the Bayes Factor (see file `model_comparison.Rmd` in the associated online repository). Bayes Factors quantify the quality of predictions of a model, averaging over the possible values of the parameters of the models (weighted by their prior probability); by averaging over the prior distribution on parameters, Bayes Factor implicitly take into account model complexity because models with more parameters will tend to have a broader prior distribution over parameters, which in effect, can water down the possible better predictions that a model with more parameters can achieve [@lee2014bayesian]. For this analysis, we treated age continuously. Table \@ref(tab:table5) lists the Bayes factors for the different model comparisons. The results show that the **integration model**, by far, outperformed all the other models. When comparing the lesioned models among each other, we see that models including the mutual exclusivity inference make better predictions compared to the *no mutual exclusivity model*.

Taken together, these analyses showed two things. First, the **integration model** makes accurate predictions about how mutual exclusivity and common ground inferences are integrated. It does so based on knowing the strength and development of each inference alone, and incorporating them into a structured probabilistic model of pragmatic reasoning. Second, models that omit one or more types of information (object specific word knowledge, speaker informativeness, common ground sensitivity) make appreciably worse predictions. This result exemplifies that children across the entire age range flexibly integrate all the available information. In the next section we ask whether there are other ways think about the process of information integration than the way formalized in the **integration model**.

@@

In this section we explore how we can best explain information integration across development. We explore alternative ways to think about information integration. The **integration model** outlined above operates via Bayesian inference in that the prior probability of a referent (a consequence of the common ground manipulation) is updated via the likelihood of hearing the utterance heard from a speaker, which is used to derive the mutual exclusivity inference. In essence, this model is a multiplicative model because the posterior probability of each referent given the utterance is proportional to the product of the likelihood of a speaker saying that utterance and the prior probability of that referent.

Information sources need not be integrated in a Bayesian manner. Rather than be part of the same pragmatic reasoning architecture, the common ground and mutual exclusitivity inferences we consider could be computed separately and then combined. A listener could integrate these two independent inferences in an additive manner, weighting the sources of information by some ratio $\phi$. $\phi$ is then a bias to prefer one type of inference relative to the other. We formalize this alternative hypothesis as a  *mixture model*.

The **integration** and **mixture** models have potentially different implications for developmental change.
The **integration model** assumes that the process by which information is integrated remains constant across development. What changes is children's semantic knowledge, their expectations about speaker informativeness, and their sensitivity to common ground, but the way these information sources are combined remains the same across age. The **mixture model** alternatively posits that there is some bias or preference for one information source, and this bias could change across development. This **developmental mixture model**  makes the same assumptions about developmental change in the individual information sources as the **integration model** but, in addition, assumes that the way that the mutual exclusivity and the common ground inference are combined with one another changes over time. This model is structurally identical to the **mixture model** but the mixture parameter $\phi$ is a function of age (i.e. represented by an intercept and a slope).

In this section, we make use of all the available data to arbitrate between these alternative models (a fully Bayesian analysis). As before, Experiments 1 and 2 directly constrain the parameters governing semantic knowledge, speaker informativeness, and the prior distribution over referents (via common ground; Expt. 2). Now, in addition, we incorporate the data from Experiment 3 to additionally constrain these parameters as well as inform the mixture parameter for the mixture model (see Figure \@ref(fig:bda)).

@@

<!-- \begin{equation}  -->
$$P^{mixture}_{L_1}(r, \mathcal{L}|u) = \phi \cdot P_{S_1}(u|r_{t}, \mathcal{L}) + (1-\phi) P( \mathcal{L})P(r)$$
<!-- \end{equation}  -->

@@

For this model, we make the mixture parameter $\phi$ a function of age and estimate the intercept and slope that yield the best model predictions compared to the data from Experiment 3. The model code can be found in the associated online repository (file: `webppl/explanation_mixture_model.wppl`). Figure \@ref(fig:fig10)B visualizes the developmental trajectory of the mixture parameter. Based on this model, the common ground inference seems to decrease in importance compared to the mutual exclusivity inference with age.

@@

We compared models based on correlations and Bayes factors. Figure \@ref(fig:fig11) shows correlations between model predictions and the data, each binned by year and object. Even though model predictions and data are closely aligned for all models, the **integration model** shows the highest correlation in all age groups. Next we directly compared models based using Bayes factors (see Table \@ref(tab:table6)). We also included the prediction **integration model** into this analysis.

Perhaps unsurprisingly, we see that further constraining the parameters for semantic knowledge, speaker informativeness and common ground sensitivity by the data from Experiment 3 greatly improves the model fit (comparison: integration (explanation) vs integration (prediction)). We also see that the explanation **integration model** provides, by far, the best fit to the data compared to the two **mixture models**. Interestingly, the prediction **integration model** also had a better fit, even though its parameters were not constrained by the data from Experiment 3. When comparing the two **mixture models** directly, we see that an age sensitive mixture parameter did not result in a substantially better fit.

@@

Next, we tested the predictive power of this model. That is, we asked how well the model would predict the data of Experiment 3, when only knowing the developmental trajectories for mutual exclusivity (based on Experiment 1) and common ground (Experiment 2). We found a very close alignment of the model predictions and the data across the entire age range. Furthermore, the **integration model** provided a better fit to the data compared to a number of lesioned models, which selectively omitted one type of information. This suggests that children flexibly integrate all available information.

In the final section, we studied which process best explained children's information integration. We compared the **integration model** to a **mixture model** which assumed that children are biased towards one type of inference. We found that the **integration model** better explained the data compared to these alternative model. In sum, we found that children's integration of semantic and pragmatic information during word learning is best described as a form of Bayesian social inference.

>>>>

```{r, include = F}
library(tidyverse)
library(rwebppl)
library(brms)
library(coda)
library(ggthemes)
library(readxl)
library(ggpubr)
library(stringr)
library(matrixStats)
library(data.table)
library(langcog)
library(ggimage)
library(BayesFactor)
library(pdftools)

@@

$$
P_{L}(r \mid u)\propto P_{S}(u \mid r) \cdot P(r \mid \rho_i) \\ 
P_{S}(u \mid r)\propto P(r\mid u) ^{\alpha_i} \\
P(r\mid u) \propto \mathcal{L}(u, r \mid \theta_{ij})
$$

@@

The model with each data-analytic parameter represented explicitly in $\{, \}s$ is:

$$
P_{L}(r \mid u; \{\rho_i, \alpha_i\, \theta_{ij}\})\propto P_{S}(u \mid r; \{\alpha_i, \theta_{ij}\}) \cdot P(r \mid \rho_i) \\ 
P_{S}(u \mid r; \{\alpha_i\, \theta_{ij}\})\propto P(r \mid u; \{\theta_{ij}\}) ^{\alpha_i} \\
P(r \mid u; \{\theta_{ij}\}) \propto \mathcal{L}(u, r \mid \theta_{ij})
$$

@@

We compare this model to three alternative, "lesioned" models that lack certain components of the full integration model.

@@

A marginally less-complex model than the full integration model uses the same model architecture, taking in speaker informativeness and common ground sensitivity, but omits semantic knowledge that is specific to the familiar objects (i.e., uses only general semantic knowledge). We described above that the parameters for semantic knowledge are fitted via a hierarchical regression (mixed effects) model. In this model, there is an overall developmental trajectory for semantic knowledge (main effect) and then there is object-specific variation around this trajectory (random effects). The **no word knowledge model** takes in the overall trajectory for semantic knowledge---represented by $\mu^\theta_0$ and $\mu^\theta_1$---but ignores object-specific variation. Formally, the semantic knowledge for each item is the same: $\theta_{ij} = \mu^\theta_0 +  i \cdot \mu^\theta_1$. That is, the model assumes a listener whose mutual exclusivity inference does not vary depending on the particular familiar object but only depends on the average semantic knowledge.

@@

We consider a model that ignores common ground information, $\rho_i$. This model takes in object specific semantic knowledge and speaker informativeness butuses a prior distribution over objects that is constant across alignment conditions and uniform (e.g., [0.5, 0.5]). This model corresponds to a listener who only focuses on the mutual exclusivity inference and ignores the common ground manipulation. As a consequence, the listener does not differentiate between the two common ground alignment conditions.

@@

Finally, we consider a model that focuses only on common ground information and ignores the identity of the objects on the tables as well as any inferences their semantic knowledge of the familiar objects license. The predictions of this model correspond to the prior distribution over objects $P(r \mid \rho_i)$.

@@

We compared the models mentioned above in two ways. First, we used correlations between model predictions and the data. For this analysis, we binned the model predictions and the data by age in years and by the type of familiar object. Figure \@ref(fig:fig6) visualizes the correlation between model predictions and the data for all models. The results shows a very high correlation between the predictions of the **integration model** and the data in all age groups indicating that the model accurately captures the variation in the data. Correlations for the **integration model** were also higher compared to the other models considered. The correlation increased from 2- to 3-year-olds but then again dropped for 4-year-olds. The mis-match for the oldest age group is probably a consequence of the model making very extreme predictions in the congruent condition. This results from the fact that 4-year-olds show very high performance both the mututal exclusivity and common ground tasks. When combined in the model, the two inferences amplify one another because we assume no cost of integration. We maintain no cost to integration because positing integration cost introduces a free parameter that could only be estimated based on the integration data of Experiment 3 itself (i.e., such a model would not make precise *a priori* predictions).

We additionally compared models based on the marginal likelihood of the data under each model -- the likelihood of the data averaging over ("marginalizing over") the prior distribution on parameters; the pair-wise ratio of marginal likelihoods for two models is known as the Bayes Factor (see file `model_comparison.Rmd` in the associated online repository). Bayes Factors quantify the quality of predictions of a model, averaging over the possible values of the parameters of the models (weighted by the prior probabilities of those parameter values); by averaging over the prior distribution on parameters, Bayes Factor implicitly take into account model complexity because models with more parameters will tend to have a broader prior distribution over parameters, which in effect, can water down the potential gains in predictive accuracy that a model with more parameters can achieve [@lee2014bayesian]. For this analysis, we treated age continuously. Table \@ref(tab:table5) lists the Bayes factors for the different model comparisons. The results show that the **integration model**, by far, outperformed all the other models. When comparing the lesioned models among each other, we see that models including the mutual exclusivity inference make better predictions compared to the *no mutual exclusivity model*.

Taken together, these analyses showed two things. First, the **integration model** makes accurate predictions about how mutual exclusivity and common ground inferences are integrated. It does so based on knowing the strength and development of each inference alone, and incorporating them into a structured probabilistic model of pragmatic reasoning. Second, models that omit one or more types of information (object specific word knowledge, speaker informativeness, common ground sensitivity) make appreciably worse predictions. This result exemplifies that children across the entire age range flexibly integrate all the available information. In the next section we ask investigate alternative formulations of the process of information integration.

@@

In this section we explore how to best explain information integration across development. We explore alternative ways to think about information integration. The **integration model** outlined above operates via Bayesian inference in that the prior probability of a referent (a consequence of the common ground manipulation) is updated via the likelihood of hearing the utterance heard from a speaker, which is used to derive the mutual exclusivity inference. In essence, this model is a multiplicative model because the posterior probability of each referent given the utterance is proportional to the product of the likelihood of a speaker saying that utterance and the prior probability of that referent.

Information sources need not be integrated in a Bayesian manner as part of the same pragmatic reasoning architecture. Instead, the common ground and mutual exclusitivity inferences we consider could be computed separately and combined in a more heuristic fashion. A listener could integrate these two independent inferences in an additive manner, weighting the sources of information by some ratio $\phi$. $\phi$ is then a bias to prefer one type of inference relative to the other. We formalize this alternative hypothe sis as a  *mixture model*.

The **integration** and **mixture** models have potentially different implications for developmental change.
The **integration model** assumes that the process by which information is integrated remains constant across development. What changes is children's sensitivity to different cues: semantic knowledge, their expectations about speaker informativeness, and their sensitivity to common ground. The **mixture model** alternatively posits that there is some bias or preference for one information source, and this bias could change across development. This **developmental mixture model**  makes the same assumptions about developmental change in sensitivities to individual information sources as the **integration model** but, in addition, assumes that the way that the mutual exclusivity and the common ground inference are combined changes over time. This model is structurally identical to the **mixture model** but the mixture parameter $\phi$ is a function of age, which we model via a logistic-linear model: $\phi_i = \text{logistic}(\beta^\phi_0 + i \cdot \beta^\phi_1)$, whose parameters we infer via Bayesian inference.

In this section, we make use of all the available data to arbitrate between these alternative models (a fully Bayesian analysis). As before, Experiments 1 and 2 directly constrain the parameters governing semantic knowledge, speaker informativeness, and the prior distribution over referents (via common ground; Expt. 2). Now, in addition, we incorporate the data from Experiment 3 to additionally constrain these parameters as well as inform the mixture parameter for the mixture models (see Figure \@ref(fig:bda)).

@@

<!-- \begin{equation}  -->
$$
P^{mixture}_{L}(r \mid u; \{\phi, \rho_i, \alpha_i, \theta_{ij} \}) = \phi \cdot P_{ME}(r \mid u; \{\alpha_i, \theta_{ij}\})  + (1-\phi) \cdot P(r \mid \rho_i) \\
P_{ME}(r \mid u; \{\alpha_i, \theta_{ij}\}) \propto P_{S}(u \mid r; \{\alpha_i, \theta_{ij}\})
$$
where $P_{ME}(r \mid u; \{\alpha_i, \theta_{ij}\})$ represents a mutual-exclusivity inference from a listener with a uniform prior on referents. $P_{S}(u \mid r_{t}$ is the same as in Eq. S above.
<!-- \end{equation}  -->

@@

For this model, we make the mixture parameter $\phi$ as a logistic-linear function of age-- $\phi_i = \text{logistic}(\beta^\phi_0 + i \cdot \beta^\phi_1)$ --- and estimate the intercept and slope that yield the best model predictions compared to the data from Experiment 3. The mixture model is then:
$$
P^{dev\_mixture}_{L}(r \mid u; \{\phi_i, \rho_i, \alpha_i, \theta_{ij} \}) = \phi_i \cdot P_{ME}(r \mid u; \{\alpha_i, \theta_{ij}\})  + (1-\phi_i) \cdot P(r \mid \rho_i) \\
$$
The model code can be found in the associated online repository (file: `webppl/explanation_mixture_model.wppl`). Figure \@ref(fig:fig10)B visualizes the developmental trajectory of the mixture parameter. Based on this model, the common ground inference seems to decrease in importance compared to the mutual exclusivity inference with age.

@@

As before, we compared models based on correlations and Bayes factors. Figure \@ref(fig:fig11) shows correlations between model predictions and the data, each binned by year and by object. Even though both mixture models have more free parameters then the *integration model*, the **integration model** shows the highest correlation in all age groups, though model predictions and data are closely aligned for all models. Next we directly compared models based using Bayes factors (see Table \@ref(tab:table6)). We also included the prediction **integration model** into this analysis.

Perhaps unsurprisingly, we see that informing the parameters for semantic knowledge, speaker informativeness and common ground sensitivity by the data from Experiment 3 greatly improves the model fit (comparison: integration (explanation) vs integration (prediction)). We also see that the explanation **integration model** provides, by far, the best fit to the data compared to the two **mixture models**, even though the integration model has fewer free parameters than the mixture model. Interestingly, the prediction **integration model** also had a better fit, even though its parameters were not constrained by the data from Experiment 3. When comparing the two **mixture models** directly, we see that an age sensitive mixture parameter did not result in a substantially better fit, providing additional evidence that the integration process itself is not changing across development.

@@

Next, we tested the predictive power of this model. That is, we asked how well the model would predict the data of Experiment 3, when only knowing the developmental trajectories for mutual exclusivity (based on Experiment 1) and common ground (Experiment 2). We found a very close alignment of the model predictions and the data across the entire age range. Furthermore, the **integration model** provided a better fit to the data compared to a number of lesioned models, which selectively omitted one type of information. These results suggest that children flexibly integrate all available information.

In the final section, we studied which process best explained children's information integration. We compared the **integration model** to a **mixture model** which assumed that children are biased towards one type of inference, and a **developmental mixture model** that assumed that the bias could change across development. We found that the **integration model** better explained the data compared to both of these mixture models, even though the mixture models had more free parameters. In sum, we found that children's integration of semantic and pragmatic information during word learning is best described as a form of Bayesian social inference.
